Requantization of time-dependent mean field for pairing collective motion in superfluid nuclei

### 2018-12-17 thesis/thesis.tex
thesis

%%% PAGE DIMENSIONS
%\usepackage{geometry} % to change the page dimensions
\usepackage[top=25truemm,bottom=20truemm,left=20truemm,right=20truemm]{geometry}
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margins=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

@@

In most of nuclei, the structure of low-lying excited $0^+$ states, which had been called as $\beta$ vibration or two-phonon excitation, are complicate and mysterious from plenty of recent experimental data. The common suggestion from those experimentalists is that the pairing correlation plays an important role in such excited $0^+$ states because they found strange cross sections of two-nucleon transfer reactions. From theoretical viewpoint, 
%excited $0^+$ states in superfluid nuclei correspond to not only quadrupole vibrational mode, the elementary mode of excitation of quadrupole collective motions, but also pairing vibrational mode, the elementary mode of excitation of pairing collective motions. 
the competition between quadrupole correlation and pairing correlation causes complicate structure of low-lying excited $0^+$ states. 
We aim to understand such excited $0^+$ states based on time-dependent mean-field (TDMF) theory. Because quantum fluctuation is large in low-lying excitation region, theoretical framework in large-amplitude regime is necessary. However, the theoretical framework to describe pairing collective motions in large-amplitude regime is not developed. The nature of pairing collective motions in low-lying excitation region is still unclear.

We consider the theoretical framework to describe large-amplitude collective motions associated with nuclear pairing. In general, the low-lying excited states can be obtained in two steps: (1) solve TDMF equation to obtain classical trajectories; (2) quantize such classical trajectories to obtain the excited states. The step 2 is called ``requantization''. In this thesis, we focus on step 2 and discuss the requantization method in detail. To simplify the system, apart from realistic nuclear system, we consider the pairing collective motions in multi-level pairing model. We divide the pairing system into two classes, integrable system and non-integrable system. The former one corresponds to two-level system and the later one corresponds to the system more than three-level. Because the description of requantization is different between the two, we consider the requantization methods for each case, respectively.

In integrable system, the requantization of TDMF is particular feasible because we can easily find the periodic orbits. We apply three different requantization methods, canonical quantization, Fourier decomposition of the time-dependent observables, and stationary phase approximation (SPA) to the path integral. Comparing each requantization method with exact solution in various two-level system, respectively, SPA turns out to be most accurate between the three and can describe the excited states quantitatively, especially for the calculation of two-particle transition strength. Because the requantized wave functions from SPA are given in terms of the microscopic degrees of freedom, the transition matrix elements can be calculated directly.

Actually, SPA cannot be applied to non-integrable system straightforwardly because finding the periodic orbits is extremely difficult task. We propose a new theoretical approach which enables SPA to apply to non-integrable system, namely, combining SPA with adiabatic self-consistent collective coordinate (ASCC) method (ASCC+SPA). The concept is that we employ ASCC to extract an integrable subspace from TDMF phase space, and apply SPA to obtain the excited states. The detailed derivation is shown in the text.
From the application to the low-lying excited $0^+$ states in three-level system and Pb isotopes system, we find ASCC+SPA gives a reasonable description.

We also examine the performance of the pairing collective model which phenomenologically assumes pairing gap parameter and gauge angle as collective coordinates. We check the one-to-one correspondence between the collective coordinates from TDMF dynamics itself and those in collective model. We find the collective coordinate describing pairing vibration has no one-to-one correspondence between the two, in both integrable and non-integrable system. The applicability of the collective model is limited for the nuclear pairing.

We summarize the achievements in this thesis. We find a reasonable requantization method, SPA, from the study in integrable system. Using SPA, we construct a theoretical framework of ASCC+SPA to requantize the pairing collective motions in non-integrable system. Also, we demonstrate that to describe pairing vibration properly, it is necessary to employ the collective coordinate extracted from TDMF degrees of freedom itself, instead of the phenomenological collective coordinate, pairing gap parameter. With the development of computational power, we think 
ASCC+SPA is promising tool to elucidate the structure of mysterious low-lying excited $0^+$ states in nuclei.

@@

Mean-field model is standard tool to describe the structure of atomic nuclei. The idea of mean-field model is based on the primary theoretical framework of shell model introduced by Mayer and Jensen in 1949. Mean field picture indicates that the nucleon can be described in single-particle motion in the mean field potential created by the whole nucleons in the nucleus. Although the many-body correlation between the nucleons is extremely complicate, the single-particle motion picture can well reproduce the basic properties (e.g. radius, binding energy, separation energy, magic number) in most of nuclide. %%The most popular mean field potential is Woods-Saxon potential with spin-orbit coupling. Because

@@

Of course, single-particle motion picture cannot describe the nuclear system perfectly. To describe the nuclear structure more precisely, the residual interaction, the deviation between single-particle interaction and many body correlation, needs to be taken account. In 1950s, nuclear physicists found one of the most important residual interaction, pairing correlation. One of the clear evidences for pairing correlation detected from experiment is the separation energy difference between odd nuclei and even nuclei. Fig. \ref{S_n} shows the separation energies for Sn, Sb, and Te isotopes as a function of neutron number. The isotopes in even neutron number have about $2\sim3$ MeV larger separation energies than the isotopes in odd neutron number.
It is odd-even staggering (OES) effect, which is detected in whole mass region of nuclei (Fig. \ref{gap}). In this figure, $2\Delta$ corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei. The approximate value is $2\Delta \approx 24A^{-1/2}$ MeV. It indicates a two-body attractive interaction should be included based on single-particle motion picture. In addition, if we examine the spin-parity of the ground states, we can find all of the even-even nuclei are $J^{\pi}=0^+$ states and most of the odd mass number nuclei have the same spin-parity as their single-particle orbital which the last nucleon fulfills in.
To explain these phenomena, it is natural to consider that the ground state of even-even nuclei consist of coupled $J^{\pi}=0^+$ pairs and the ground state of odd mass number nuclei consist of coupled $J^{\pi}=0^+$ pairs except one unpaired nucleon in last single-particle orbit. The two-body attractive interaction which couples two identical nucleons into $J^{\pi}=0^+$ state is called pairing correlation.

@@

Because pairing correlation is two-body interaction, a theoretical framework beyond the mean-field model seems to be necessary. However, there is a tricky approximated theory, BCS theory, enable us to describe pairing correlation within mean-field model. In BCS theory, the ground state is superposed by the states with different particle number. It indicates that the ground state breaks symmetry in a space, namely U(1) gauge space. With symmetry breaking, the corresponding order parameter, pairing gap $\Delta$, has finite value. Based on BCS theory, because the energy of each particle in a $J^{\pi}=0^+$ pair is $E_k=\sqrt{(\epsilon_k-\lambda)^2+\Delta^2}$, where $\epsilon_k$ is single particle energy, $\lambda$ is Fermi energy, the lowest single-particle excited state corresponds to one-pair broken state with minimum excitation energy $2\Delta$ In fact, pairing gap corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei discussed in previous paragraph. As same as superconductivity in condensed matter physics, the superfluid nucleus is defined when $\Delta\neq0$ in ground state. Thus, the OES detected from experiment caused from pairing correlation can be well explained by BCS theory.

After BCS theory was firstly introduced by J. Bardeen, L. Cooper, and J. R. Schrieffer in 1957, it was immediately generalized to Hartree-Fock-Bogoliubov (HFB) theory. HFB theory is the naturally extension of Hartree-Fock (HF) theory by introducing Bogoliubov transformation. Under Bogoliubov transformation, the particles become to quasiparticles, and the mean-field potential is extended to generalized form including pairing field. After HFB theory had been applied to nuclear physics in 1958, the structure of ground states in nuclei was examined by plenty of studies. Up to now, HFB theory is the standard mean-field theory to describe nuclear structure including pairing correlation.

@@

Due to the superfluid state breaks the symmetry in U(1) gauge space, we expect that there are two types of elementary modes of excitation for pairing collective motion in nuclei. The elementary modes of excitation are Nambu-Goldstone (NG) mode, pairing rotational mode, recovering the broken symmetry in U(1) gauge space, and Higgs mode, pairing vibrational mode, fluctuating around the energy minimum point.

First, we discuss about pairing rotation. In classical picture, pairing rotation is the motion which the system rotates in the U(1) gauge space (red bi-direction arrow in Fig. \ref{pair_rot}). The motion is described by the coordinate, gauge angle $\Phi$, and the corresponding momentum, total particle number $N$. 
%It is similar to normal rotation in 3D space which described by the angles ($\theta_x, \theta_y, \theta_z$) and their conjugate angular momenta ($J_x, J_y, J_z$). 
Because the total particle number is unchanged in time evolution, pairing rotation corresponds to constant of motion. In quantum picture, if we assume the ground state of $N_0$ particle system as the ``ground state'', the ground states in $N=N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' of pairing rotational mode. The excitation energy is $E_{\rm rot}^{\rm pair}=\frac{(N-N_0)^2}{2\mathcal{J}}$, where $\mathcal{J}$ is the pairing-rotational moment of inertia. In experiment, the pairing rotational band is expected to be observed in even-even isotope chain or even-even isotone chain in open-shell nuclei. A clear observation is found in Sn isotopes shown in Fig. \ref{pair_rot}. Except a constant value and linear term with respect to mass number $A$, we can find the energies of ground states consist of a parabolic line - pairing rotational band. Recently, the systematic study for the pairing-rotational moment of inertia has been done in Ref. \cite{HN16}. The experimental moments of inertia obtained from two neutron(proton) separation energies have excellent correspondence with the theoretical moments of inertia obtained from NG mode in quasiparticle random phase approximation (QRPA) \cite{TV61} based on HFB mean-field theory, not only for the spherical nuclei, but also for the deformed nuclei. Because the pairing-rotational moments of inertia are free from ambiguities caused from odd-even system, they are supposed to be excellent indicators of pairing correlation rather than pairing gaps.

Next, we discuss about pairing vibration. We emphasize that there are two types of pairing vibrations, one which vibrates around normal ground state ($\Delta\approx 0$), and another one which vibrates around superfluid ground state ($\Delta\neq 0$). The structures of two vibrational modes are quite different. 
%The former one is well understood because the structure of the excited $0^+$ states are $2np2nh$ configurations, detected from the energy spectra around ${}^{208}$Pb. 
Here, we only focus on the later case.
In classical picture, the system vibrates around one of the degenerated energy minimum points. Depending on the intrinsic structure, various pairing vibrations are able to exist in one nucleus. 
%In general, the motion is supposed to be described by the coordinate, pairing gap $\Delta$, and its corresponding momentum\footnote{Because $\Delta$ is order parameter, it is naturally supposed to be the coordinate to describe the pairing vibration in previous studies. However, we found $\Delta$ is not proper coordinate in most of case from our study. The detail is discussed in Chapter 3 and 4.}.
In quantum picture, if we assume the ground state of $N_0$ particle system as the ``ground state'', the excited states in $N=N_0, N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' of pairing vibrational mode. The excitation energy is $E_{\rm vib}^{\rm pair}=n_{\alpha}\hbar\omega$, where $n_{\alpha}$ is the number of phonon and $\alpha$ is the label to distinguish different pairing vibrational modes. Because $2\Delta$ is the minimum energy to break one-pair, the collectivity is supposed to be strong when $E_{\rm vib}^{\rm pair}<2\Delta$. Thus, low-lying excited $0^+$ states in even-even nuclei are possible to be pairing vibrational states. Except the region around the magic number, excited $0^+$ states below $2\Delta$ exist in most of even-even nuclei. However, it is difficult to observe the signatures for pairing vibration directly from experiment. The properties of pairing vibration in nuclei are not well understood up to now.
% because the multipole correlations interfere the pairing correlation significantly

The signatures for pairing collective motion can also be observed from the two-neutron(proton) transfer reaction \cite{BHR73}, such as $(p, t)$, $(t, p)$, and $({}^3{\rm He}, n)$. The cross sections of two particle transfer reaction strongly depend on the matrix elements of $J^{\pi}=0^+$ pair-transition operators (e.g. pair-additional transition: $\sigma( P_{ad};0_{\alpha}^+(N) \to 0_{\beta}^+(N+2))\sim |\braket{N+2,\beta|\sum_k a^{\dag}_k a^{\dag}_{\bar{k}}|N,\alpha}|^2$). From theoretical prediction, the pairing correlation is known to enhance $\sigma(g.s.\to g.s.)$ and decrease $\sigma(g.s.\to 0_{ex}^+)$ significantly, hence, $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)\ll 1$ is expected in superfluid system. One example of $(p, t)$ reactions is demonstrated in Fig. \ref{pair_rot}. As mentioned in previous paragraph, the ground states of even-even Sn isotopes consist of a pairing rotational band. In such situation, we can find intraband transition is dominant with respect to interband transition. Therefore, the cross sections of two-neutron transfer reaction is consistent with theoretical prediction. In general, the ratio $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)$ for two-neutron(proton) transfer reaction are excellent indicators to describe the collectivity of pairing dynamics.

\subsection{Mysterious excited $0^+$ states and pairing dynamics}
In most of nuclei, the structure of excited $0^+$ states in a low-lying region (a few MeV excitation) is complicate and unclear \cite{HW11}. The reason is that not only pairing correlation but also deformations and multipole correlations influence the excited $0^+$ states significantly. The various correlations competing with each other construct plenty of mysterious excited $0^+$ states. We present several examples for these excited $0^+$ states observed from experiment.

After the concept of pairing vibration was introduced, it encountered the problem for the large asymmetry cross sections between $(p, t)$ and $(t, p)$ reaction in actinide region (e.g. Th, U, Pu, Cm) observed from experiment. In low-lying energy region, the cross section of $\sigma(g.s.\to 0_{ex}^+)$ is about $15\%$ of $\sigma(g.s.\to g.s.)$ for $(p, t)$ reaction \cite{MEFSS70, MEFSS72}, while the signal of $\sigma(g.s.\to 0_{ex}^+)$ is extremely weak for $(t, p)$ reaction \cite{Cas72}. To explain such phenomena, the coexistence of both oblate shape and prolate shape in one state, and the mixing of pairing correlations between oblate and prolate shape was suggested \cite{GJV71,RK72}. Another interpretation for such phenomena is using microscopic interaction, including multipole pairing correlations and quadrupole particle-hole interactions \cite{RB76}. Although the approaches are different, they reproduced the asymmetry cross sections of two-neutron transfer reaction in actinide region. These studies hinted the importance of the quadrupole deformation and interactions in low-lying excited $0^+$ states.

We can consider the excited $0^+$ states from common viewpoint, liquid drop with surface vibration. Bohr collective model, which is well-known and successful model describing the quadrupole collective motions in nuclei \cite{BM75} has been constructed based on the idea of liquid drop. The $0_2^+$ states in the model correspond to $\beta$ vibration and the first $2^+$ state in $K=2$ bands correspond to $\gamma$ vibration, in quadrupole deformed nuclei. Also, their rotational bands exist in each vibrational mode. The nuclei in rare earth region (e.g. Sm, Gd, Dy, Er, Yb, Hf, W), well deformed nuclei, had been supposed to be the typical examples of quadrupole collective motions in the early days. 
With the development for extracting $B(E2)$ values from experiment, however, the doubt whether $0_2^+$ states in rare earth region is $\beta$ vibration arose in the last decades \cite{G01, Kulp08, SS11}. They found that only a part of $B(E2)$ values related to the $0_2^+$ states agree with the predicted $B(E2)$ values in pure $\beta$ vibration picture. Moreover, for $E0$ transition, very few $0_2^+$ states satisfies the range of $\rho^2(E0)$ values which the $\beta$ vibrational picture is good. On the other hand, they detected strong cross sections $\sigma(g.s.\to 0_{2}^+)$ in both $(t, p)$ and $(p, t)$ reactions. These results suggest that $0_2^+$ states in rare earth region are not pure $\beta$ vibrational states, but some mixed states with the importance of pairing correlation.

\begin{figure}[tb]
 \begin{center}
    \includegraphics[width=100mm]{images/phonon.eps}
 \end{center}
 \begin{center}
    \includegraphics[width=100mm]{images/112Cd.eps}
 \end{center}
  \caption{$B(E2)$ values, relative to the $B(E2; 2_1^+\to 0_1^+)$ value, and energy spectra in phonon excitation picture. Upper panel: expected $B(E2)$ values for a quadrupole harmonic vibrator. Lower panel: data observed for ${}^{112}$Cd. Dashed transitions indicate an unobserved transition. Detailed information is discussed in Ref. \cite{GW10}.}
  \label{112Cd}
\end{figure}
The mystery of excited $0^+$ states exist not only in deformed nuclei, but also in spherical nuclei. A best example is the structure of $0_2^+$ or $0_3^+$ states in Cd isotopes observed in recent experiment \cite{GW10}. The lower part of Fig. \ref{112Cd} shows the excitation energy and $B(E2)$ transition strength in ${}^{112}$Cd. Comparing with the typical values for the ideal quadrupole harmonic vibration in spherical nuclei shown in the upper part of Fig. \ref{112Cd}, we can find that the observed excitation energy spectra have good coincidence with phonon excitation picture. The $0_2^+$ or $0_3^+$ states seem to can be interpreted as two-phonon or three-phonon excitation. However, the $B(E2)$ values are very different comparing with each other, especially for the no observed decay from $0_{2ph}^+$ to $2_1^+$ state. Such phenomena are also seen in other Cd isotopes. The configuration of these $0^+$ states are supposed to be not pure phonon excitation, but proton pair-excitation between $2p_{1/2}$ and $1g_{9/2}$ orbits from the hint of the ground state configurations of Ag isotopes \cite{HW11}. The pairing should plays an important role if the excited $0^+$ states are described by pair-excitation.

@@

\clearpage{\pagestyle{empty}\cleardoublepage}
\chapter{Conclusion}
\label{conclusion}

>>>>

%%% PAGE DIMENSIONS
%\usepackage{geometry} % to change the page dimensions
\usepackage[top=30truemm,bottom=25truemm,left=20truemm,right=20truemm]{geometry}
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margins=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

@@

In early days, the low-lying excited $0^+$ states in most of nuclei had been considered as $\beta$ vibration or two-phonon excitation associated with quadrupole vibrations. However, plenty of recent experimental data indicate that not only quadrupole correlations but also pairing correlations play important roles in such excited $0^+$ states. 
%From theoretical viewpoint, 
%excited $0^+$ states in superfluid nuclei correspond to not only quadrupole vibrational mode, the elementary mode of excitation of quadrupole collective motions, but also pairing vibrational mode, the elementary mode of excitation of pairing collective motions. 
The competition between quadrupole correlations and pairing correlations causes complicate and mysterious structure for the low-lying excited $0^+$ states. 
We aim to understand such excited $0^+$ states based on time-dependent mean-field (TDMF) theory. Because quantum fluctuation is large in low-lying excitation region, we need to consider the collective motion in large-amplitude regime. However, the theoretical framework to describe pairing collective motions in large-amplitude regime is not well developed. The nature of pairing collective motions in low-lying excitation region is still unclear.

We consider the theoretical framework to describe large-amplitude collective motions associated with nuclear pairing. In general, the low-lying excited states can be obtained in two steps: (1) solve TDMF equation to obtain the classical trajectories; (2) quantize such classical trajectories to obtain the excited states. The step 2 is called ``requantization''. Because the theoretical framework for the requantization is not well developed, we focus on the step 2 and discuss it in detail. 
%To simplify the system, 
Apart from realistic nuclear system, 
we consider the pairing collective motions in multi-level pairing model. We divide the pairing system into two classes, integrable system and non-integrable system. The former one corresponds to two-level system and the later one corresponds to the system more than three-level. Because the description of requantization is different between the two, we consider the requantization methods for each case, respectively.

In integrable system, the requantization of TDMF is particular feasible because we can easily find the periodic orbits. We apply three different requantization methods, canonical quantization, Fourier decomposition of the time-dependent observables, and stationary phase approximation (SPA) to the path integral. Comparing each requantization method with exact solution in various two-level system, respectively, SPA turns out to be most accurate between the three and can describe the excited states quantitatively, especially for the description of two-particle transition strength. Because the requantized wave functions from SPA are given in terms of the microscopic degrees of freedom, the transition matrix elements can be calculated directly.

Actually, SPA cannot be applied to non-integrable system straightforwardly because finding the periodic orbits is extremely difficult task. We propose a new theoretical approach which enables SPA to apply to non-integrable system, that is, combining SPA with adiabatic self-consistent collective coordinate (ASCC) method (ASCC+SPA). The concept is that we employ ASCC to extract an integrable subspace from original TDMF phase space, and apply SPA to obtain the excited states. 
%The detailed derivation is shown in the text.
After the application to the low-lying excited $0^+$ states in three-level system and Pb isotopes system, we find that ASCC+SPA can properly describe the pairing collective motions.

We also examine the performance of the pairing collective model which assumes pairing gap parameter and gauge angle as collective coordinates. We check the one-to-one correspondence between the collective coordinates from TDMF dynamics itself and those in the collective model. We find the collective coordinate describing pairing vibration has no one-to-one correspondence between the two, in both integrable and non-integrable system. The applicability of the collective model is limited for the nuclear pairing.

We summarize the achievements in this thesis. We find a reasonable requantization method, SPA, from the study in integrable system. Using SPA, we construct a theoretical framework of ASCC+SPA to requantize the pairing collective motions in non-integrable system. Also, we demonstrate that it is necessary to describe the pairing vibration in the collective coordinate extracted from TDMF degrees of freedom itself, instead of the phenomenological collective coordinate, pairing gap parameter. With the development of computational power and numerical techniques, we think 
the ASCC+SPA is the promising tool to elucidate the structure of mysterious low-lying excited $0^+$ states in nuclei, in the near future.

@@

Mean-field model is a standard tool to describe the structure of atomic nuclei. The idea of mean-field model is based on the primary theoretical framework of shell model introduced by Mayer and Jensen in 1949. Mean field picture indicates that the nucleus can be described in single-particle motion in the mean field potential created by the whole nucleons in the nucleus. Although the many-body correlations between the nucleons is extremely complicate, the single-particle motion picture can well reproduce the basic properties (e.g. radius, binding energy, separation energy, magic number) in most of nuclide. %%The most popular mean field potential is Woods-Saxon potential with spin-orbit coupling. Because

@@

Of course, single-particle motion picture cannot describe the nuclear system perfectly. To describe the nuclear structure more precisely, residual interactions need to be taken account. In 1950s, nuclear physicists found one of the most important residual interactions, pairing correlation. One of the clear evidences for pairing correlation detected from experiment is the separation energy difference between odd nuclei and even nuclei. Fig. \ref{S_n} shows the separation energies for Sn, Sb, and Te isotopes as a function of neutron number. The isotopes in even number of neutron have about $2\sim3$ MeV larger separation energies than the isotopes in odd number of neutron.
It is odd-even staggering (OES) effect, which is detected in whole mass region of nuclei (Fig. \ref{gap}). In this figure, $2\Delta$ corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei. The approximate value is $2\Delta \approx 24A^{-1/2}$ MeV. It indicates a two-body attractive interaction should be included based on single-particle motion picture. In addition, if we examine the spin-parity of the ground states, we can find all of the even-even nuclei are $J^{\pi}=0^+$ states and most of the odd mass number nuclei have the same spin-parity as their single-particle orbit which the last nucleon fulfills in.
To explain these phenomena, it is natural to consider that the ground state of even-even nuclei consist of coupled $J^{\pi}=0^+$ pairs and the ground state of odd mass number nuclei consist of coupled $J^{\pi}=0^+$ pairs except one unpaired nucleon in last single-particle orbit. The two-body attractive interaction which couples two identical nucleons into $J^{\pi}=0^+$ state is called pairing correlation.

@@

Because pairing correlation is two-body interaction, a theoretical framework beyond the mean-field model seems to be necessary. However, there is a tricky approximated theory, BCS theory, enable us to describe pairing correlation within mean-field model. In BCS theory, the states with different particle number are superposed to describe the ground state. It indicates that the ground state breaks symmetry in a space, namely U(1) gauge space. With symmetry breaking, the corresponding order parameter, pairing gap $\Delta$, has finite value. Based on BCS theory, because the energy of each particle in a $J^{\pi}=0^+$ pair is $E_k=\sqrt{(\epsilon_k-\lambda)^2+\Delta^2}$, where $\epsilon_k$ is single particle energy, $\lambda$ is Fermi energy, the lowest single-particle excited state corresponds to one-pair broken state with minimum excitation energy $2\Delta$. In fact, pairing gap corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei. As same as superconductivity in condensed matter physics, the superfluid nucleus is defined when $\Delta\neq0$ in ground state. Thus, BCS theory can explain the OES effect caused from pairing correlation.

After BCS theory was firstly introduced by J. Bardeen, L. Cooper, and J. R. Schrieffer in 1957, it was immediately generalized to Hartree-Fock-Bogoliubov (HFB) theory. HFB theory is the naturally extension of Hartree-Fock (HF) theory by introducing Bogoliubov transformation. Under Bogoliubov transformation, the particles become to quasiparticles, and the mean-field is extended to generalized form including pairing field. After HFB theory had been applied to nuclear physics in 1958, the structure of ground states in nuclei was examined by plenty of studies. Up to now, HFB theory is the standard mean-field theory to describe nuclear structure including pairing correlation.

@@

Due to the superfluid state breaks the symmetry in U(1) gauge space, there are two types of elementary modes of excitation for pairing collective motion in nuclei. The elementary modes of excitation are Nambu-Goldstone (NG) mode, pairing rotational mode, recovering the broken symmetry in U(1) gauge space, and Higgs mode, pairing vibrational mode, fluctuating around the energy minimum point.

First, we discuss the pairing rotation. In classical picture, pairing rotation is the motion which the system rotates in the U(1) gauge space (red bi-direction arrow in Fig. \ref{pair_rot}). The motion is described by the coordinate, gauge angle $\Phi$, and the corresponding momentum, total particle number $N$. 
%It is similar to normal rotation in 3D space which described by the angles ($\theta_x, \theta_y, \theta_z$) and their conjugate angular momenta ($J_x, J_y, J_z$). 
Of course, $N$ is a constant of motion. In quantum picture, if we assume the ground state of $N_0$ particle system as the ``ground state'', the ground states in $N=N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' in the pairing rotational mode. The excitation energy is $E_{\rm rot}^{\rm pair}=\frac{(N-N_0)^2}{2\mathcal{J}}$, where $\mathcal{J}$ is the pairing-rotational moment of inertia. In experiment, the pairing rotational band is expected to be observed in even-even isotope chain or even-even isotone chain in open-shell nuclei. A clear observation is found in Sn isotopes shown in Fig. \ref{pair_rot}. Except a constant value and linear term with respect to mass number $A$, we can find the energies of ground states consist of a parabolic line - pairing rotational band. Recently, the systematic study for the pairing-rotational moment of inertia has been done in Ref. \cite{HN16}. The experimental moments of inertia obtained from two neutron(proton) separation energies have excellent correspondence with the theoretical moments of inertia obtained from NG mode in quasiparticle random phase approximation (QRPA) \cite{TV61}, not only for the spherical nuclei, but also for the deformed nuclei. Because the pairing-rotational moments of inertia are free from ambiguities caused from odd-even system, they are supposed to be excellent indicators of pairing correlation rather than pairing gaps.

Next, we discuss the pairing vibration. We emphasize that there are two types of pairing vibrations, one vibrating around a normal ground state ($\Delta\approx 0$), and the other vibrating around superfluid ground state ($\Delta\neq 0$). The properties of the two are quite different. 
%The former one is well understood because the structure of the excited $0^+$ states are $2np2nh$ configurations, detected from the energy spectra around ${}^{208}$Pb. 
Here, we only focus on the later case.
In classical picture, the system vibrates around one of the degenerated energy minimum points. Depending on the intrinsic structure, various pairing vibrations are able to exist in one nucleus. 
%In general, the motion is supposed to be described by the coordinate, pairing gap $\Delta$, and its corresponding momentum\footnote{Because $\Delta$ is order parameter, it is naturally supposed to be the coordinate to describe the pairing vibration in previous studies. However, we found $\Delta$ is not proper coordinate in most of case from our study. The detail is discussed in Chapter 3 and 4.}.
In quantum picture, if we assume the ground state of $N_0$ particle system as the ``ground state'', the excited states in $N=N_0, N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' of pairing vibrational mode. The excitation energy is $E_{\rm vib}^{\rm pair}=n_{\alpha}\hbar\omega$, where $n_{\alpha}$ is the number of phonon and $\alpha$ is the label to distinguish different pairing vibrational modes. Because $2\Delta$ is the minimum energy to break one-pair, the collectivity is dominant when $E_{\rm vib}^{\rm pair}<2\Delta$. Thus, low-lying excited $0^+$ states in even-even nuclei have strong collectivity associated with pairing vibration. Except the region around the magic number, excited $0^+$ states below $2\Delta$ exist in most of even-even nuclei. However, it is difficult to observe the signatures for pairing vibration directly from experiment. The properties of pairing vibration in nuclei are not well understood up to now.
% because the multipole correlations interfere the pairing correlation significantly

The signatures for pairing collective motion can also be observed from the two-neutron(proton) transfer reaction \cite{BHR73}, such as $(p, t)$, $(t, p)$, and $({}^3{\rm He}, n)$. The cross sections of two particle transfer reaction strongly depend on the matrix elements of $J^{\pi}=0^+$ pair-transition operators (e.g. pair-additional transition: $\sigma( P_{ad};0_{\alpha}^+(N) \to 0_{\beta}^+(N+2))\sim |\braket{N+2,\beta|\sum_k a^{\dag}_k a^{\dag}_{\bar{k}}|N,\alpha}|^2$). From theoretical prediction, the pairing correlation is known to enhance $\sigma(g.s.\to g.s.)$ and decrease $\sigma(g.s.\to 0_{ex}^+)$ significantly, hence, $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)\ll 1$ is expected in superfluid system. One example of $(p, t)$ reactions is demonstrated in Fig. \ref{pair_rot}. As mentioned in previous paragraph, the ground states of even-even Sn isotopes consist of a pairing rotational band. In such situation, the intraband transition is dominant with respect to the interband transition. Therefore, the cross sections of two-neutron transfer reaction is consistent with theoretical prediction. In general, the ratio $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)$ for two-neutron(proton) transfer reaction are excellent indicators to describe the collectivity of pairing dynamics.

\subsection{Mysterious excited $0^+$ states and pairing dynamics}
In most of nuclei, the structure of excited $0^+$ states in a low-lying region (a few MeV excitation) is complicate and unclear \cite{HW11}. The reason is that not only pairing correlations but also deformations and multipole correlations influence the excited $0^+$ states significantly. The various correlations competing with each other construct plenty of mysterious excited $0^+$ states. We present several examples for these excited $0^+$ states observed from experiment.

After the concept of pairing vibration was introduced, it encountered the problem for the large asymmetry cross sections between $(p, t)$ and $(t, p)$ reaction in actinide region (e.g. Th, U, Pu, Cm) observed from experiment. In low-lying energy region, the cross section of $\sigma(g.s.\to 0_{ex}^+)$ is about $15\%$ of $\sigma(g.s.\to g.s.)$ for $(p, t)$ reaction \cite{MEFSS70, MEFSS72}, while the signal of $\sigma(g.s.\to 0_{ex}^+)$ is extremely weak for $(t, p)$ reaction \cite{Cas72}. To explain such phenomena, the coexistence of both oblate shape and prolate shape in one state, and the mixing of pairing correlations between oblate and prolate shape was suggested \cite{GJV71,RK72}. Another interpretation for such phenomena is using microscopic interaction, including multipole pairing correlations and quadrupole particle-hole interactions \cite{RB76}. Although the approaches between these studies are different, they reproduced the asymmetry cross sections of two-neutron transfer reaction in actinide region. These studies hinted the importance of the quadrupole correlations in low-lying excited $0^+$ states.

We can consider the excited $0^+$ states from common viewpoint, liquid drop with surface vibration. Bohr collective model, which is well-known and successful model describing the quadrupole collective motions in nuclei \cite{BM75} has been constructed based on the idea of liquid drop. The $0_2^+$ states in the model correspond to $\beta$ vibration and the first $2^+$ state in $K=2$ bands correspond to $\gamma$ vibration, in quadrupole deformed nuclei. Also, their rotational bands exist in each vibrational mode. The nuclei in rare earth region (e.g. Sm, Gd, Dy, Er, Yb, Hf, W), well deformed nuclei, had been supposed to be the typical examples of quadrupole collective motions in the early days. 
With the development for extracting $B(E2)$ values from experiment, however, the doubt whether $0_2^+$ states in rare earth region is $\beta$ vibration arose in the last decades \cite{G01, Kulp08, SS11}. They found that only a part of $B(E2)$ values related to the $0_2^+$ states agree with the predicted $B(E2)$ values in pure $\beta$ vibration picture. Moreover, for $E0$ transition, very few $0_2^+$ states satisfies the range of $\rho^2(E0)$ values which the $\beta$ vibrational picture predicts. On the other hand, they detected strong cross sections $\sigma(g.s.\to 0_{2}^+)$ in both $(t, p)$ and $(p, t)$ reactions. These results suggest that $\ket{0_2^+}$ in rare earth region are not pure $\beta$ vibrational states, but some states mixing with pairing correlations.

\begin{figure}[tb]
 \begin{center}
    \includegraphics[width=100mm]{images/phonon.eps}
 \end{center}
 \begin{center}
    \includegraphics[width=100mm]{images/112Cd.eps}
 \end{center}
  \caption{$B(E2)$ values, relative to the $B(E2; 2_1^+\to 0_1^+)$ value, and energy spectra in phonon excitation picture. Upper panel: expected $B(E2)$ values for a quadrupole harmonic vibrator. Lower panel: data observed for ${}^{112}$Cd. Dashed transitions indicate an unobserved transition. Detailed information is discussed in Ref. \cite{GW10}.}
  \label{112Cd}
\end{figure}
The mystery of excited $0^+$ states exist not only in deformed nuclei, but also in spherical nuclei. A best example is the structure of $0_2^+$ or $0_3^+$ states in Cd isotopes observed in recent experiment \cite{GW10}. The lower part of Fig. \ref{112Cd} shows the excitation energy and $B(E2)$ transition strength in ${}^{112}$Cd. Comparing with the typical values for the ideal quadrupole harmonic vibration in spherical nuclei shown in the upper part of Fig. \ref{112Cd}, we can find that the observed excitation energy spectra have good coincidence with phonon excitation picture. The $0_2^+$ or $0_3^+$ states seem to be interpreted as two-phonon or three-phonon excitation. However, the $B(E2)$ values are quite different between the two, especially for the no observed decay from $0_{2ph}^+$ to $2_1^+$ state. Such phenomena are also seen in other Cd isotopes. The configuration of these $0^+$ states are supposed to be not pure phonon excitation, but proton pair-excitation between $2p_{1/2}$ and $1g_{9/2}$ orbits from the hint of the ground state configurations of Ag isotopes \cite{HW11}. The pairing should play an important role if the excited $0^+$ states are described by pair-excitation.

@@

\clearpage{\pagestyle{empty}\cleardoublepage}
\chapter{Conclusion and perspective}
\label{conclusion}
